### 2014-11-22 papers/paper2.tex

\chapter*{Massivley Multiplayer Online Role-Playing Games: The Past, Present, and Future}

  \begin{itemize}
    \item {\BF MMORPG:} Massivley Multiplayer Online Role-Playing Games
    \item 
  \end{itemize}