### 2014-11-23 papers/paper3.tex
Added templates for the rest of the papers

\chapter*{Persuasive Games: Bringing Computer Entertainment Back to the Real World}