### 2014-11-23 papers/paper8.tex

\chapter*{From Visual Simulation to Virtual Reality to Games}