### 2014-11-24 papers/paper2.tex
changed chapter headers and added content to paper7

\chapter*{2) Massivley Multiplayer Online Role-Playing Games: The Past, Present, and Future}

  \begin{itemize}
    \item {\bf MMORPG:} Massivley Multiplayer Online Role-Playing Games
  \end{itemize}