### 2014-11-24 papers/paper3.tex

\chapter*{3) Persuasive Games: Bringing Computer Entertainment Back to the Real World}