### 2014-11-24 papers/paper6.tex

\chapter*{6) GameFlow: A Model for Evaluating Player Enjoyment in Games}