### 2014-11-24 papers/paper8.tex

\chapter*{8) From Visual Simulation to Virtual Reality to Games}